# Knowing the Data
The Titanic dataset is available for download from Kaggle. Then you can tell RStudio to read your dataset by doing the following:
```{r}
Titanic_train = read.csv("data/train.csv")
```

Let’s look at a data summary.
```{r}
summary(Titanic_train)
```

I discovered something strange in the Survived and Sex section. So I used the sapply() function to determine the data class.
```{r}
sapply(Titanic_train, class)
```
Then I altered the classes of Survive, Sex, SibSp, and Parch by doing the following:
```{r}
Titanic_train$Survived = as.factor(Titanic_train$Survived)
Titanic_train$Sex = as.factor(Titanic_train$Sex)
Titanic_train$SibSp = as.numeric(Titanic_train$SibSp)
Titanic_train$Parch = as.numeric(Titanic_train$Parch)
```

Let’s go over the data class again with the sapply() function. Also, double-check the data summary.

The Titanic dataset has 12 variables. As a result, I classified variables into two types.
1. Continuous variables: Age, SibSp, Parch, and Fare.
2. Categorical variables: PassengerId, Survived, Pclass, Sex, Ticket, Cabin, and Embarked.

# Cleaning the Data
To count the number of NA values in a column, I used the is.na() function.
```{r}
sum(is.na(Titanic_train))
```
Then I decided to eliminate the row that has NA value by following:
```{r}
Titanic_train <- Titanic_train[!(is.na(Titanic_train$Age)),]
```

# Visualizing the Data
First and foremost, you must install a data visualization package. I chose to install the ggplot2 package and ran the library() function.
```{r}
library(ggplot2)
```

1. Survival by Sex
```{r}
ggplot(data=Titanic_train, aes(Sex, fill = factor(Survived))) + 
  geom_bar(stat = "count", position = "fill") + 
  xlab("Sex") + scale_fill_discrete(name = "Survived") + 
  ylab("%Passengers") + ggtitle("Proportion of Passengers by Sex") + 
  scale_y_continuous(labels = function(x) paste0(x*100, "%"))
```

2. Survival by Age
```{r}
ggplot(data=Titanic_train, aes(Age, fill = factor(Survived))) +
  geom_density(aes(fill = as.factor(Survived)), alpha = 0.5) +
  xlab("Age") +
  ylab("Density") +
  labs(fill = "Survived") +
  ggtitle("Age Distribution of Passengers")
ggplot(data=Titanic_train, aes(Survived, Age)) +
  geom_boxplot(alpha = 0.2) +
  scale_y_continuous(trans = "log2") +
  geom_point(show.legend = FALSE) + 
  geom_jitter()
```

3. Survival by Pclass
```{r}
ggplot(data=Titanic_train, aes(Pclass, fill = factor(Survived))) + 
  geom_bar(stat = "count", position = "fill") + 
  xlab("Pclass") + scale_fill_discrete(name = "Survived") +
  ylab("%Passengers") +
  ggtitle("Proportion of Passengers by Pclass") +
  scale_y_continuous(labels = function(x) paste0(x*100, "%"))
```

4. Survival by SibSp
```{r}
ggplot(data=Titanic_train, aes(SibSp, fill = factor(Survived))) +
  geom_density(aes(fill = as.factor(Survived)), alpha = 0.5) +
  xlab("SibSp") +
  ylab("Density") +
  labs(fill = "Survived") +
  ggtitle("SibSp Distribution of Passengers")
```

5. Survival by Parch
```{r}
ggplot(data=Titanic_train, aes(Parch, fill = factor(Survived))) +
  geom_density(aes(fill = as.factor(Survived)), alpha = 0.5) +
  xlab("Parch") +
  ylab("Density") +
  labs(fill = "Survived") +
  ggtitle("Parch Distribution of Passengers")
```

6. Survival by SibSp+Parch
```{r}
ggplot(data=Titanic_train, aes(SibSp+Parch, fill = factor(Survived))) +
  geom_density(aes(fill = as.factor(Survived)), alpha = 0.5) +
  xlab("SibSp+Parch") +
  ylab("Density") +
  labs(fill = "Survived") +
  ggtitle("SibSp+Parch Distribution of Passengers")
```

7. Survival by Fare
```{r}
ggplot(data=Titanic_train, aes(Fare, fill = factor(Survived))) +
  geom_density(aes(fill = as.factor(Survived)), alpha = 0.5) +
  xlab("Fare") +
  ylab("Density") +
  labs(fill = "Survived") +
  ggtitle("Fare Distribution of Passengers")
ggplot(data=Titanic_train, aes(Survived, Fare)) +
  geom_boxplot(alpha = 0.2) +
  scale_y_continuous(trans = "log2") +
  geom_point(show.legend = FALSE) + 
  geom_jitter()
```

8. Survival by Embarked
```{r}
ggplot(data=Titanic_train, aes(Embarked, fill = factor(Survived))) + 
  geom_bar(stat = "count", position = "dodge") + xlab("Embarked") + 
  scale_fill_discrete(name = "Survived") + 
  ggtitle("Survival by Port of Embarkation")
```

9. Table of summary statistics
I opted to utilize the gtsummary package to summarize data statistics.
```{r}
library(gtsummary)
Titanic_train <- Titanic_train %>% select(Survived, Sex, Age, Pclass, SibSp, Parch, Fare, Embarked)
Titanic_train %>% tbl_summary(by = Survived, type = c(SibSp, Parch) ~ "continuous", 
statistic = list(all_continuous() ~ "{mean} ({sd})",
all_categorical() ~ "{n} / {N} ({p}%)"))
```

